Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-11-07 demystify.tex
updates

\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{\textsc{The Simpsons}}
If you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!
In the US, the common types of funding for PhDs are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
RA is paid by a prof. for you to do their research. TA is paid by the dept. for you to help with teaching. Finally, fellowship is independent funding that can come from a school, a company, or an organization. \autoref{tab:funding} summarizes the differences.
Note that funding is typically more available for PhD students than
MS.

@@

\begin{commentbox}[Vu:]
  TA and RA at Mason have similar benefits in tuition waiving and insurance.  The college and department will set the rate for a 9-month graduate assistant stipend.  TA, which is paid by the department, will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: close to graduate.}) and the prof.
  \tcblower
  Having health insurance is required at many US universities.  Do not assume that you're young and healthy and ignore insurance (\autoref{sec:cultural-misc}).  At Mason, and most CS PhD programs, your GTA or GRA comes with full insurance. In fact, at Mason your spouse/children will get a significant discount rate for health insurance.  So you will never have to worry much about health issues for you or your family here.
\end{commentbox}

@@

\subsection{Application Materials}

>>>>

\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{\textsc{The Simpsons}}

If you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!
In the US, the common types of funding for PhDs are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
RA is paid by a prof. for you to do their research. TA is paid by the dept. for you to help with teaching. Finally, fellowship is independent funding that can come from a school, a company, or an organization.

\autoref{tab:funding} summarizes the differences.
Note that funding is typically more available for PhD students than MS (\autoref{sec:ms-funding}).

@@

\begin{commentbox}[Vu:]
  TA and RA at Mason have similar benefits in tuition waiving and insurance.  The college and department will set the rate for a 9-month graduate assistant stipend.  TA, which is paid by the department, will likely be that amount but RA might be higher depending on  the prof and the stage of the student (1st year vs ABD\footnote{All but dissertation: close to graduate.}).\\

  At Mason, CS TA/RA is paid the highest.  First, CS is under the College of Engineering and Computing (CEC), which already has a higher TA stipend than other colleges.  Second, CS is the largest department in the college, and has the higher TA stipend than what CEC suggests.  Finally, it is a tradition that CS faculty pay their RAs more than the department's TA rate.  So a winning combination for CS PhD students.

  \tcblower
  Having health insurance is required at many US universities.  Do not assume that you're young and healthy and ignore insurance (\autoref{sec:cultural-misc}).  At Mason, and most CS PhD programs, your GTA or GRA comes with full insurance. In fact, at Mason your spouse/children will get a significant discount rate for health insurance.  So you will never have to worry much about health issues for you or your family here.
\end{commentbox}

@@

\subsection{Application Materials}\label{sec:ms-application}
